---
title: "Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load:

```{r}
library(dplyr)
library(tidyr)
library(rstackdeque)
library(skimr)

load("../../subset_data.Rdat")
```

The `skimr` package makes nice table summaries, but the customization syntax is... ugh. This defines a skimming function where factor columns are summarized with more characters and top entries.

```{r}
myskimmer <- skim_with(factor = sfl(top_counts = ~top_counts(., max_char = 50, max_levels = 8)))
```

## condition_era.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$condition_era.csv)
```

## condition_occurrence.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$condition_occurrence.csv)
```


## device_exposure.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$device_exposure.csv)
```

## drug_era.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$drug_era.csv)
```

## drug_exposure.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$drug_exposure.csv)
```


## goldstandard.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$goldstandard.csv)
```


## measurement.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$measurement.csv)
```

## observation_period.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$observation_period.csv)
```


## observation.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$observation.csv)
```



## person.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$person.csv)
```


## procedure_occurrence.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$procedure_occurrence.csv)
```

## visit_occurrence.csv

```{r}
#skim_format(.levels = list(max_char = 36))
myskimmer(data$visit_occurrence.csv)
```

